``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
require('plyr')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
source('~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/uniformity.R')

data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric',
                             'factor','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','character','character'))

# no idea why these numbers are messed up, but this fix seems to solve it
data$Speaker.Voronoiness = ifelse(data$Speaker.Voronoiness > 1, 
                                  yes = data$Speaker.Voronoiness / 1000, 
                                  no = data$Speaker.Voronoiness)
data$Hearer.Voronoiness = ifelse(data$Hearer.Voronoiness > 1, 
                                 yes = data$Hearer.Voronoiness / 1000, 
                                 no = data$Hearer.Voronoiness)

ds = data[,c(1,4,5,7,8,11:15,17)] # only relevant data
dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
```

Uniformity and diversity in evolutionary outcomes
=================================================

Within each triplet of independent parameter values, how much variation do we see in the evolving lanugages? It turns out that a side-effect of impairment is that it unifies evolutionary outcomes. To see this, we can plot the difference between the maximmal and the minimal value of each property for each triplet of independent variables.


```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
MinMaxProperty = ddply(dm, .(Number.of.states, Impairment, Tolerance, variable), 
                      summarise, 
                      MaxMinDiff = max(value) - min(value),
                      MinProperty = min(value), MaxProperty = max(value), MeanProperty = mean(value))

pd <- position_dodge(.01)
sp = ggplot(MinMaxProperty, aes(x=Impairment, y=MaxMinDiff, colour=factor(Number.of.states))) + 
     geom_point(position = pd)

sp + facet_grid(variable ~ Tolerance, scales="free")
```

We can then look at the maximal difference among all properties. (This works because all relevant dependent measures scale inbetween 0 and 1.)

```{r}
diffTrials = ddply(MinMaxProperty,
                   .(Number.of.states, Impairment, Tolerance), 
                  summarise, 
                  diffTrials = max(MaxMinDiff) <=1e-5,
                  maxDiffTrials = max(MaxMinDiff) )

ggplot(diffTrials, aes(x=Impairment, y=maxDiffTrials, )) + 
     geom_point() + facet_grid(Number.of.states ~ Tolerance, scales="free")
```

We see clearly that, all else equal (e.g. fixed tolerance and number of states), impairment 'unifies' evolutionary outcomes. When impairment is at least 0.1, outcomes are all almost identical (in terms of the recorded properties).

Notice that this 'uniformity' is different from just convexity vs. non-convexity!

```{r message=FALSE, warning=FALSE, fig.width=5, fig.height=5}

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
d = subset(dms, dms$variable == "Speaker.Convex.Cat")

pd <- position_dodge(.01)
sp = ggplot(d, aes(x=Impairment, y=1-value )) + 
     geom_point(position = pd)

sp + facet_grid(Number.of.states ~ Tolerance, scales="free")

```

There are strategies whose speaker strategies are all convex (e.g. for ns = 6, tolerance = 0.3 and impairment = 0.05), but whose other properties differ. Here are two strategies from that class of independent variable values that differ maximally in terms of expected utility.

```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

d = subset(data, Number.of.states == 6 & Tolerance == 0.3 &
             Impairment == 0.05 & Speaker.Convex.Cat == 1)

show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))

```

In contrast, if we look at a "uniform" class of independent parameter values, such as with ns = 50, tolerance = 0.1, impairment = 0.1, then the strategies whose EU are minimal and or maximal in that class don't differ optically at all (modulo arbitrariness of message use).

```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

d = subset(data, Number.of.states == 50 & Tolerance == 0.1 &
             Impairment == 0.1)

show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))

```

What this suggests is that the differences in some of the groups might be due to non-evenly splitting the interval. Impairment leads to even splits and therefore to uniformity of evolutionary outcomes.

Mean distance between speaker strategies 
----------------------------------------

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
  show(sPlot)
```

Putting it all together 
-----------------------


```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
  show(combPlot)
```

